---
title: "Homework 9"
output: github_document
---
Homework Project Part 1: Final Project Prospectus

 For our final project, we aim to create an R package that can conduct analyses 
of growing microbial culture and bacterial population relative abundance data. 
Regarding data obtained from microbial cultures, we will enable our package to 
integrate excel data tables, statistically process the encoded data, and 
generate scatter/line plots that will display data in time vs. cell count 
format. The data used for this will include data that Frank has collected in 
the Buchan lab on the bacterial species Sulfitobacter sp. CB-D.

 Additionally, we also aim to create a package that can perform analyses of many 
different types of abundance data. In the future, we hope to use this package to
analyze data that Trystan will collect from the McKinney lab studying 
concentrations and bioaccumulation of heavy metals in aquatic and terrestrial 
organisms. This data is not time-specific, so it is important that we also 
create a package where abundance is not reliant on a time variable. 



```{r setup, include=FALSE}
library(tidyverse)
```
As always, the first step is to load the required libraries. For this you will
only need the library tidyverse.

Homework Problem #1

The goal here is develop a function that can calculate the variance of a numeric
vector. This function should also auto-print the result or store it as a 
variable.

```{r}
numeric_vector <- c(1, 2, 2, 2, 3, 4, 5, 5, 5)

variance <- function(x) {
  (1 / (length(x) - 1)) * sum( (x - mean(x) ) ^2)
  }

skewness <- function(y) {
  ((1 / (length(y) - 2)) * sum((y - mean(y) ) ^3)) / (variance(y) ^(3/2))
  }
```

With these new functions, "variance()" and "skewness", any vector input into 
them will auto-print the respective mathematical value of the vector. For 
instance, when inputing our vector, "numeric_vector", into the function 
"variance()", we receive the following:

```{r}
variance(numeric_vector)
```
Likewise, if we input "numeric_vector" into the function "skewness()", we 
receive the following:
```{r}
skewness(numeric_vector)
```

The above code accomplishes the desired task. First, a numeric vector is 
created, stored as "numeric_vector". Functions are then developed to calculate 
both variance and skewness. These functions are then applied to the numeric 
vector and the results are auto-printed.

Homework Problem #2

The goal of this problem is to develop a function that will determine if the 
value NA is present in the same position in two different vectors. Firstly, we 
will create some vectors to use.

```{r}
na_vector_1 <- c(1, 2, 3, NA, 5, 6, NA)

na_vector_2 <- c(1, NA, 3, NA, 5, 6, 7)

na_vector_3 <- c(1, NA, 3, NA, 5, 6, NA)
```

After developing these vectors, we need to create a function that will look for
the value "NA" and confirm its presence in two vectors. This is accomplished via
the function below.

```{r}
both_na <- function(x,y) { 
  length(intersect( which(is.na(x)), which(is.na(y))))
  }
```

This function works by completing several tasks. First, the "is.na()" function 
determines the positions in the data set that have "NA" stored. These values are
returned using boolean logic.

```{r}
is.na(na_vector_1)
```
The "which()" function specifically asks, which of these values are "TRUE". 
Reading out the entire function, which(is.na(na_vector_1)), it is asking which 
positions in the vector "na_vector_1" store the value "NA".
```{r}
which(is.na(na_vector_1))
```
In this case, the positions 4 and 7 have the value "NA". Next, the function 
"intersect()" is used to ask, which of the vectors observed have values that 
intersect. In this case, we can look at the functions below.

```{r}
which(is.na(na_vector_1))
which(is.na(na_vector_2))
```
We see that in the reported values, both datasets report that their fourth 
values are "NA". By using the intersect value was see this as well.

```{r}
intersect( which(is.na(na_vector_1)), which(is.na(na_vector_2)))
```
As the does not ask for WHICH positions in the vectors both contain the value 
"NA", but rather how many, we only need to determine the length of this
returned vector. This can be done using the function "length()".

```{r}
length(intersect( which(is.na(na_vector_1)), which(is.na(na_vector_2))))
```
As we can see, the two vectors only both had "NA" in the fourth position, so 
only one position in both vectors stored "NA". The function below completes all 
of these steps at once.

```{r}
both_na(na_vector_1, na_vector_2)
```
We can also look at this using the third vector assigned earlier. Simply by 
observing the vectors we see that positions 4 and 7 both store in "NA" in 
vectors "na_vector_1" and "na_vector_3". Therefore, the result here is 2.

```{r}
both_na(na_vector_1, na_vector_3)
```


Homework Problem #3

The purpose of this assignment is to use a function to create a usable graph
from a datframe. This is roughly done by the code below.

First, the data of interest must be stored as a data frame. In this case, we 
store "Practice_Dataset.csv" as a date frame, labeled "colony_counts". This
dataset contains data containing cell densities of the bacteria 
Sulfitobacter sp. CB-D. 

```{r}
colony_counts <- read.csv("data/Practice_Dataset.csv")
```

Next, a function is created that will manipulate the data and generate a useful
plot.

```{r}
scatter_plot <- function(z) {
  long_colony_counts <- pivot_longer(z,
               cols = starts_with("T"),
               names_to = "Time_Points", 
               values_to = "Cell_Density")
  
  ggplot(long_colony_counts, aes(x = Time_Points, y = Cell_Density))+
    geom_point()
}
```

The function above first takes the desired dataset and converts into a long 
format. It takes all columns that with T, in this case Timepoint columns, and
stores them in all in a new column labeled "Time_Points". It also takes the 
cell density values associated with each time point and stores them in another
new column termed, "Cell_Density". This can be seen below.

```{r}
long_colony_counts <- pivot_longer(colony_counts,
          cols = starts_with("T"),
          names_to = "Time_Points", 
          values_to = "Cell_Density")

print(long_colony_counts)
```
After storing this data set, the rest of the function generates a table, 
plotting "Time_Points" vs "Cell_Density" as a scatter plot.

```{r}
scatter_plot(colony_counts)
```
While this graph needs some work, it completes the basics of this problem. It 
can also now be easily relabeled. The next steps here will be to average trial 
replicates, so as to automate the process in R instead of doing it by hand, and
color code the points present.


Homework Problem #4

The goal of this problem is to design a function that can multiply numeric 
vectors by 2, sort character vectors alphabetically, and report an error if 
neither of these are done. 

The function required needs to be able to judge vectors based on their class. 
This can be done by using a series of if-else statements, shown in the function 
below. 

```{r}
vector_sorting <- function(z) {if (class(z) == "numeric") {
   z*2
  } else if (class(z) == "character"){
  sort(z)
  } else {
  warning("Error: Data provided is neither a numeric nor character vector.")
  }
}
```
The code above uses an initial if statement, stating if this vector is numeric, 
multiply it by 2. In the case of the "na_vector_1", "na_vector_2", and 
"na_vector_3" vectors used earlier, this statement will return the following.
```{r}
vector_sorting(na_vector_1)
```
```{r}
vector_sorting(na_vector_2)
```
```{r}
vector_sorting(na_vector_3)
```
If the provided data is not numeric, it will move to the next "else-if"
function. The next function asks if the class of the vector is a character. If
it is, sort it alphabetically. We can test this by storing a new character 
vector and then applying it to the "vector_sorting" function.

```{r}
animals <- c('dog', 'cat', 'zebra', 'fish', 'antelope', 'buffalo', 'chicken')

vector_sorting(animals)
```
We can see that the function indeed sorted the character vector alphabetically.
However, if the data provided is also not a character vector, the function will 
go to the statement, in this case an else statement. Now, the function will 
return an error message, stating that the provided data are neither numeric or 
character classed vectors. This informs the user that they should confirm the 
data they are inputting is accurate/appropriate. By testing the function on 
the previously stored "colony_counts" data frame we see this error report.
```{r}
vector_sorting(colony_counts)
```
This function successfully multiplies numeric vectors by 2, alphabetically 
sorts character vectors, and reports an error when an inappropriate dataset is 
provided.

Homework Problem #5

The purpose of problem 5 is to generate the mean of a vector. If it is 
successful it returns the mean, while if unsuccessful it returns an informative
error message. This is accomplished using the code below.

```{r}
mean_calculation <- function(x) {if (class(x) == "numeric"){
  return (mean(x))
  } else {
    warning("Error: Mean calculation was unsuccessful. Please ensure that the
    dataset of interest is strictly numeric.") 
  }
  
}
```

Essentially this function will return the mean of any numeric vector/data, but 
will attempt to calculate the mean of objects it regularly would not. In these
cases it reports an error. This can be seen through the examples below.

In this example, we use a previously created numeric vector, where the mean is 
quickly calculated.

```{r}
mean_calculation(numeric_vector)
```
In this example, we attempt to calculate the mean of the previously generated
"animals" vector. As this is not a numeric vector, the function will return an 
error.

```{r}
mean_calculation(animals)
```


Homework Problem #6

The purpose of problem #6 is to determine how long certain processes take that
complete the same function. The goal is to compare a simple summation of two 
1-million random element numeric vectors through a sum() function as well as 
two iterations of for loops, one storing the sum in an assigned third vector and
one not storing the third vector. 

First we need to create our 1-million random element numeric vectors. This is 
done by the commands below.

```{r}
one_million_numbers <- sample(1000000)
one_more_million_numbers <- sample(1000000)
```


The most basic summation is established below.

```{r}
one_million_sums <- (one_million_numbers + one_more_million_numbers)
glimpse(one_million_sums)
```
The system.time() command compares the length it takes to complete this command.

```{r}
system.time(one_million_sums <- (one_million_numbers + one_more_million_numbers))
```
As we can see this is very brief. Next we can create a for-loop that adds all of
these without storing the vector.

```{r}
system.time(for (i in seq_along(one_million_numbers)) {
  (one_million_numbers[i] + one_more_million_numbers[i])
})
```
In this summation, the required time is much longer. However, this is without 
actually storing the result. If we wish to store the result, the function below
can be used.

```{r}
for_loop_sum <- c()

system.time(for (i in seq_along(one_million_numbers)) {
  for_loop_sum[i] <- (one_million_numbers[i] + one_more_million_numbers[i])
})
```
As we can see here, the required time is over double that of when the values are
not stored. This show that different methods to do the same thing can require
vastly different amounts of time and resources.