---
title: "Homework 10"
output: github_document
---

Linear Models

As always, our very first step is to load the libraries we will be using. In 
this instance, we will only be using 'tidyverse'.

```{r}
library(tidyverse)
```

After loading our libraries, we can start by answering the first question. This 
problem asks us to create a linear model of price as a function of the carat of
a diamond from the 'diamonds' dataset. This can be done using the command below.

```{r}
diamonds_data <- diamonds

carat_price <- lm(price ~ carat, data = diamonds_data)
```

First we stored the diamonds dataset as our own object, then we stored a new 
object that is the linear model [lm()] of price as a function of carat values 
from the diamonds_data object.

From this value we, will want the "Intercept" and "carat" values. In a normal
formula, these are the "b" and "m" values in y = mx + b. We can store these 
values below as their own value set.

```{r}
carat_price_coefficients <- carat_price$coefficients
print(carat_price_coefficients)
```
We can now solve for y which is our "price" value, given that a single carat is 
used. So "price" = mx + b, or "price" = 7756.426(carat) + -2256.361. The formula
below calculates this for us.

```{r}
carat_cost_equation <- function(x,y) {
  (x - (y [1]))/(y [2])
}

carat_value <- 1

carat_cost_equation(carat_value, carat_price_coefficients)
```
This value of $0.29 is how much a diamond should cost based solely on the carat.
Obviously more goes into this calculation than just the weight, as 30 cents for 
a diamond is ridiculously low. Using other methods we can look at the role of 
other values along with carats on diamond price.

We can take into account two variables using the code below, color and carat, 
opposed to just the carat.

```{r}
carat_and_color_price <- lm(price ~ carat + color, data = diamonds_data)
```

After this, we can store the coefficients we desire once again.
```{r}
carat_and_color_price_coefficients <- carat_and_color_price$coefficients
print(carat_and_color_price_coefficients)
```

We now see that several values come into play now. However, we once again will 
be using the first two and one other value, dependent on the color being 
utilized. We can calculate the value of a single carat and color L using the 
formula below.

```{r}
carat_color_cost_equation <- function(x,y) {
  (x - (y [1]))/(y [2] + y[3])
}

carat_color_cost_equation(carat_value, carat_and_color_price_coefficients)
```
So, the value of a diamond per carat is $0.42. This obviously off, but it is
the value address by the linear model. If you notice previously, the coefficient
values listed were not the actual colors listed in diamonds. This can be seen 
below.

```{r}
print(summary(diamonds_data$color))
```
As you can see here, the colors are labeled D through J, not L, Q, C or any 
other color listed in the previous coefficients. This is due to the data being
an "ordered" class opposed to a "character" class. We can store the colors as
character data via the code below.

```{r}
carat_and_chr_color_price <- lm(price ~ carat + as.character(color), diamonds_data)
```

We once again store the coefficients as their own values.
```{r}
carat_and_chr_color_price_coefficients <- carat_and_chr_color_price$coefficients
print(carat_and_chr_color_price_coefficients)
```
We now see the color values are listed appropriately as opposed to before and 
that the listed color values are much different than before. They also appear to 
be more well ordered, especially as D is the highest grade of color, then E, F,
G, etc. The number decreasing in value makes sense here in that regard, whereas
before there was no obvious pattern regarding the color values. 

If we use the carat_color_cost_equation from before, we can also see the prioe 
values have changed.

```{r}
carat_color_cost_equation(carat_value, carat_and_chr_color_price_coefficients)
```
We can see this value is much closer to the original value, compared to when 
the data was in an ordered format. Now we can plot all of the lines with the 
data below to determine which line fits the data best.

```{r}
regression_comparison <- ggplot(diamonds_data, aes(carat, price)) +
  geom_point(size = 1, colour = 'grey') +
  geom_abline(
    aes(intercept = carat_price$coefficients[1], 
        slope = carat_price$coefficients[2],
        colour = "Carat")) +
  geom_abline(
    aes(intercept = carat_and_color_price$coefficients[1], 
        slope = (carat_and_color_price$coefficients[2] + 
                   carat_and_color_price$coefficients[3]),
        colour = "Carat and Color")) +
  geom_abline(
    aes(intercept = carat_and_chr_color_price$coefficients[1], 
    slope = (carat_and_chr_color_price$coefficients[2] + 
             carat_and_chr_color_price$coefficients[3]),
    colour = "Carat and Color(chr)"))+
  ggtitle(label = "Diamond Regression Line Comparison") +
  xlab("Diamond Weight (Carats)") +
  ylab("Diamond Price (US Dollars)")
  
print(regression_comparison)
```
By viewing these trends, we can see that the "Carat" and "Carat and Color" 
models fit the data the best. Distinguishing which of these is actually the line
of best fit would be difficult, but as they are so similar either should be fine
in this case.


Objects

Currently, we already have our required libraries stored from earlier, so no 
need to enter anymore library commands here. The first step thing we want to do
is to generate a new class. Firstly, we'll introduce a scenario. Say we have an
individual who works at a veterinarian's office and is storing new clients pet 
information. We can first load the data below into our chart.

```{r}
animal_information <- as.data.frame(read.csv("data/Pet_Info.csv"))
print(animal_information)
```

As you can see, we have a variety of different pets. In this case, if several
file formats have similar structures, it may be useful to generate a new class
to distinguish this data type for others.

```{r}
class_creation <- function(df, user) {
  if (! class(df) == "data.frame") {
    warning("Error: Data provided is not in a data.frame format.")
  }
  else if (! class(user) == "character"){
    warning("Error: User provided is not given as a character value.")
  }
  else { attr(df, "user") <- user
  structure(df, class = c("pet_description", "data.frame"))
  }
}
```

In the command above we do just that, shifting this "data.frame" class to a 
new structured class termed "pet_description". Note that the code specifically
ensures the original file format (df) is a "data.frame" and that the user is a
character string. Otherwise warning messages will be returned. Assuming those 
conditions are met, the final query informs the function to create a new class
called "pet_description" with the same structure as a "data.frame".

Next, we can verify this data works using the code below.

```{r}
new_class <- class_creation(animal_information, user = "John Smith")
```

This command does several things. First is calls the function "class_creation" 
that we just generated, and tells it to look at the "animal_information" data
frame. Next, it tells the class to store a user, in this case "John Smith".
When we observe the class of this "new_class" object, we see that it has been
altered from the original data.frame class.

```{r}
class(new_class)
```
As we can see, the title for this class is "pet_description" but it mimics the
formatting of a data.frame class. Now we can focus on utilizing a bit more of 
the information entered before by designing a summary command. If you remember 
before the attribute (attr) added to the initial class design function included
a "user" variable. In this case, the user can just be referred to as the one
entering the new veterinarian client information. 

```{r}
new_summary.pet_description <- function(df) {
  print(paste0("Data Filed by: ", attr(df, "user")))
  df_summ <- summary.data.frame(df)
  df_summ
}
```

By entering the "summary.pet_description" command below, we now can relay the data as well as an initial statement of who entered the data.

```{r}
new_summary.pet_description(new_class)
```
Items like this can prove useful in clarifying pertinent information about a 
dataset, in this case who entered the data. Perhaps this would allow for easy
troubleshooting, telling anyone to talk to John Smith and if any inconsistencies are found in the information he filed.
