---
title: "Homework 5"
output: github_document
---
First load the libraries necessary for the following functions. These are the
packages "nycflights13" and "tidyverse".
```{r libraries, message=FALSE}
library(nycflights13)
library(tidyverse)
```
The command below allows for the "visualization" the documentation of flights.

```{r visualization, message=FALSE}
?nycflights13::flights
```

Next, we can use the glimpse() function to view the "flights" data frame.
```{r}
glimpse(flights)
```

If we wish to see the number of total flights that were delayed by a minimum
of two hours, we can enter the command below. "120" is used as the delayed
departure time is recorded in minutes.

```{r}
a <- filter(flights, dep_delay >= 120)
print(a)
```
This results indicates that of the original 336,776 flights, 9,888 of them 
had a late departure time of at least two hours. 

It may be deemed a nuisance trying to sort through both carrier abbreviations
and flight numbers, so we can combine the two via the command below, into a 
new column labeled flight_number.

```{r}
b <- mutate(flights, flight_number = (paste(carrier,flight)))
print(b)
```
```{r}
d <- summarise(b, flight_number,year,month,day,dep_time,sched_dep_time, dep_delay,arr_time,sched_arr_time,arr_delay, carrier,flight, tailnum, origin, dest,air_time, distance, hour, minute, time_hour)
print(d)
```

Next we need to group the flights by the "carrier" and "tailnum" variables from 
the previously created data frame. This can be done using the command below.

```{r}
e <- group_by(d,carrier,tailnum)
print(e)
```

After creating this data frame above, we can summarize the results and identify 
how many times a plane with "x" tailnum flew for a specific carrier. This is 
done via the command below.

```{r}
f <- group_by(summarise(e,tailnum,n(),.groups='drop'))
print(f)
```






