---
title: "Homework 5"
output: github_document
---
First load the libraries necessary for the following functions. These are the
packages "nycflights13" and "tidyverse".
```{r libraries, message=FALSE}
library(nycflights13)
library(tidyverse)
```
The command below allows for the "visualization" the documentation of flights.

```{r visualization, message=FALSE}
?nycflights13::flights
```

Next, we can use the glimpse() function to view the "flights" data frame.
```{r}
glimpse(flights)
```

If we wish to see the number of total flights that were delayed by a minimum
of two hours, we can enter the command below. "120" is used as the delayed
departure time is recorded in minutes.

```{r}
two_hr_late <- filter(flights, dep_delay >= 120)
print(two_hr_late)
```
This results indicates that of the original 336,776 flights, 9,888 of them 
had a late departure time of at least two hours. 

It may be deemed a nuisance trying to sort through both carrier abbreviations
and flight numbers, so we can combine the two via the command below, into a 
new column labeled flight_number.

```{r}
flight_number <- flights %>%
  mutate(flight_number = (paste(carrier, flight)))
print(flight_number)
```
Next, if we wanted, we can actually reorder the statement to put the
"flight_number" column in the front. Just a personal preference.

```{r}
flight_number_reordered <- summarise(flight_number, flight_number, year, month,day, dep_time, sched_dep_time, dep_delay, arr_time, sched_arr_time, arr_delay, carrier, flight, tailnum, origin, dest,air_time, distance, hour, minute, time_hour)
print(flight_number_reordered)
```

Lastly, we want to determine how many times each plane flew for a certain 
carrier. We can do this by looking at our previous data set, then telling the
script to group the data based on the 'tailnum', or plane identifier, with the
'carrier' they flew for. Then by summarizing the number of times those two
are found identical on the same flight, we can have it relay that value,
n(), as a variable called 'planes.for.carrier'.

```{r}
planes_for_carrier <- flight_number_reordered %>%
  group_by(tailnum, carrier) %>%
  summarise(planes.for.carrier = n(), .groups='keep')
print(planes_for_carrier)
```




