---
title: "Homework 8"
output: github_document
---
Firstly, we need to load the libraries required for this assignment. All steps
will require the 'tidyverse' library while the last step will require 
'nycflights13'.

---
```{r}
library(tidyverse)
library(nycflights13)
```

Homework Step 1.
Generate a list containing three named objects. The code to do this is listed
below, with the objects 'Frank', 'Pizza', and '23', each named 'Name', 'Favorite
Food', and 'Age' respectively.

```{r}
hw_08_list <- list( Name = "Frank", 
                    Favorite_Food = "Pizza", 
                    Age = 23)
glimpse(hw_08_list)
```

The next goal is to call each of the elements in this list by different calling
mechanisms. As you can see, we call the first item on the list by it's name, 
then by [[ ]] and [] syntax respectively for the last two items on the list.

```{r}
hw_08_list["Name"]

hw_08_list[[2]]

hw_08_list[3]

```
For the next step, we need to view the library 'mtcars' 's 'mpg' column in two 
different ways. The first method we have is simply calling mpg using the single
bracket format.
`
```{r}
mtcars["mpg"]
```

We can also call 'mtcars' 's 'mpg' column by calling the column number. In this
case, we call column 1 of mtcars.
```{r}
mtcars[1]
```

The next step is to introduce a column into the mtcars dataset for "displacement
per cylinder". We can do that via the commands below, where we tell R to add 
store 'mtcars' as a new dataframe called 'mtcars.data' and then add the column
called 'displacement.per.cylinder' to it, which is the result of dividing 
the "disp" column by the "cyl" column.
```{r}
mtcars.data <- mtcars
mtcars.data["displacement.per.cylinder"] <- mtcars$disp / mtcars$cyl

print(mtcars.data)
```


By printing the dataframe, we notice that the 'displacement.per.cylinder'
column is present and we see that the values are present.

We can also address this issue using "tidyverse". Below, we tell R to, once 
again, store mtcars into our data. This time, we create a new dataframe, termed 'tidyverse_disp_per_cyl', that is mutated by the addition of a column that has 
the values of displacement divided by the number of cylinders added to it. We 
then tell this new object to only store the selected column and remove the rest
of the mtcars data.
```{r}
mtcars <- mtcars
tidyverse_disp_per_cyl <- mtcars %>%
  mutate(disp_per_cyl = (paste(disp / cyl))) %>%
  select(disp_per_cyl)
  
print(tidyverse_disp_per_cyl)
```
Via this method, we store the actual value into this dataframe. However, this 
dataframe has been stored as a character class. The significance can be seen
through the removal of "0" characters. As these were non-meaningful, when 
storing the data as a character object, it removed those characters.

Homework 8.2: Factors


```{r}
num.vec <- c(3,4,1,5)

num.fac <- as.factor(num.vec)

num.vec.v2 <- as.numeric(num.fac)
```

With the command "num.vec <- c(3,4,1,5)", you can see the output, listed below,
shows the vector stored is a series of numbers "3 4 1 5".

```{r}
glimpse(num.vec)
```
Now we convert this vector to a factor using the command 
"num.fac <- as.factor(num.vec)". The output we receive is listed below.

```{r}
glimpse(num.fac)
```
You can see now that the values returned consist of the characters "1", "3", 
"4", and "5". The order these are listed in is 2, 3, 1, 4, so while "1" is the
first character returned here, it is actually the third number in the list. The
first number in the list is the second returned value, so "3", followed by the 
third value, "4", then "1" and "5". So 2, 3, 1, 4 designates the list's order, 
while 1, 3, 4, 5, are the "alphabetical" character order of those returned 
values.

Lastly, when we return the factor to a numeric vector format, the content is
altered drastically. You can see this through the command below.

```{r}
glimpse(num.vec.v2)
```
As you can see, instead of taking the original content, it has taken the 
"alphabetical" order of those numbers instead, and returned them in this vector
conversion. All information regarding the original numbers is lost, we now only
know that the third number was the smallest, followed by the first, second, and
fourth.



Homework 8.3: Exploratory Data Analysis

This assignment requests for the exploration of data. In this case, the 
'nycflights13' dataset was chosen for analysis.

8.3.1: Delayed Flights
The first analysis that was conducted to determine the percentage of flights 
from each carrier that were at least one hour later. First, like mentioned 
before, the libraries required for this part of the homework are 'tidyverse' and 
'nycflights13'

```{r}
library(tidyverse)
library(nycflights13)
```


The next command sent is to create a new dataframe, called 'carrier_delays' that 
is derived from the 'nycflights13:flights' data set. This data will then be 
grouped by whether theflights were at least 60 minutes late and by carrier. The 
results of the grouping will be summarised in a column titled 
'delayed.flight.per.carrier'.

```{r}
carrier_delays <- flights %>%
  group_by(dep_delay >= 60, carrier) %>%
  summarise(delayed.flights.per.carrier = n(), .groups='keep')

print(carrier_delays)
```

Next, we need to divide the column containing the values for 
'delayed.flight.per.carrier' into their own unique columns based on the various
row values (e.g. 'FALSE, 'TRUE', 'NA'). To do this, we first generate a new 
object, termed 'wide_carrier_delays', that is created by having R pivot the 
data, taking column names from the 'dep_delay >= 60' column, and the values from
the 'delayed.flights.per.carrier column. We also choose to rename this data to 
make it more clear what they are testing. Column names are changed from 'FALSE' 
and 'TRUE' to 'Not.One.Hour.Late' and 'One.Hour.Late', respectively.'NA' was 
also changed to 'Unknown', as the data for these flights are not listed. This 
data will not be used in further analysis as it may skew results.

```{r}
wide_carrier_delays <- carrier_delays %>% 
      pivot_wider(names_from = 'dep_delay >= 60', values_from = 
                'delayed.flights.per.carrier') %>%
      rename(Not.One.Hour.Late = 'FALSE', One.Hour.Late = 'TRUE', 
             Unknown = 'NA')

print(wide_carrier_delays)
```


After entering this command, we can see that our data is now nicely sorted by
whether or not the flights were late, as well as by the flight carriers. Now
we need to add in some simple mathematics to determine the percentage of flights
that were late. This is done by generating a new object, termed 
'wide_carrier_delays_percentage' derived from the 'wide_carrier_delays' dataset 
previously created. This new dataset is generated by the addition of two new 
columns. Firstly, we generated one column listing the values of flights that 
were and were not one hour late, by summing the columns 'Not.One.Hour.Late' and
'One.Hour.Late', returned as 'Total.Flights'. We then add a second column which
will give us the percentage of late flights. This column, termed 
'Percent.delayed' is created by dividing the 'One.Hour.Late' flights by 
'Total.Flights' and multiplying the value by 100. This returns a percentage of 
flights that were late, by carrier.

```{r}
wide_carrier_delays_percentage <- wide_carrier_delays %>%
  mutate(Total.Flights = Not.One.Hour.Late + One.Hour.Late) %>%
  mutate(Percent.Delayed = (One.Hour.Late / Total.Flights)*100)

print(wide_carrier_delays_percentage)
```

Lastly, we can graph the data by generating a bar plot that consists of the 
individual carriers on the x-axis and the percent of delayed flights on the 
y-axis. We can also give our graph a title "Percentage of Flights Delayed by
1-Hour by Carrier".

```{r}
Percent_delayed_plot <- ggplot(wide_carrier_delays_percentage, 
                               aes(x = carrier, y =  Percent.Delayed))+
                  ggtitle(label= "Percentage of Flights Delayed by 1-Hour
                          by Carrier")+
                  geom_bar(stat = "identity")

print(Percent_delayed_plot)
```

The full length of code for the above commands is listed below.

```{r}
library(tidyverse)
library(nycflights13)

carrier_delays <- flights %>%
  group_by(dep_delay >= 60, carrier) %>%
  summarise(delayed.flights.per.carrier = n(), .groups='keep')


wide_carrier_delays <- carrier_delays %>% 
      pivot_wider(names_from = 'dep_delay >= 60', values_from = 
                'delayed.flights.per.carrier') %>%
      rename(Not.One.Hour.Late = 'FALSE', One.Hour.Late = 'TRUE', 
             Unknown = 'NA')

wide_carrier_delays_percentage <- wide_carrier_delays %>%
  mutate(Total.Flights = Not.One.Hour.Late + One.Hour.Late) %>%
  mutate(Percent.Delayed = (One.Hour.Late / Total.Flights)*100)

Percent_delayed_plot <- ggplot(wide_carrier_delays_percentage, 
                               aes(x = carrier, y =  Percent.Delayed))+
                  ggtitle(label= "Percentage of Flights Delayed by 1-Hour
                          by Carrier")+
                  geom_bar(stat = "identity")
```


8.3.2: Early-Bird Flights

The next data analysis conducted was performed to identify if different carriers
showed varying biases towards early day flights, specifically before 8 AM.

The first step is to once again load the required libraries.

```{r}
library(tidyverse)
library(nycflights13)
```

Next, we generate a dataframe labeled 'early_bird_flights' from the 
'nycflights13::flights' data. The data taken is immediately grouped by scheduled
departure time, 'dep_time', being less than 801, or 8:01 AM, as well as by 
'carrier'. Data acquired is kept in a new column termed, 
'early.bird.flights.per.carrier'.

```{r}
early_bird_flights <- flights %>%
  group_by(dep_time < 801, carrier) %>%
  summarise(early.bird.flights.per.carrier = n(), .groups='keep')

print(early_bird_flights)
```

After establishing this dataset, like previously done, we will need to pivot the
data on the desired column. We therefore take the column 'dep_time < 801', and
pivot the data on it, generating columns with names from the values in column
'dep_time < 801' and values from the integers in column 
'early.bird.flights.per.carrier'. Once again, we rename the columns listed to 
match the description as performed in the first data analysis, rather than using
'FALSE', 'TRUE', and 'NA'.

```{r}
wide_early_bird_flights <- early_bird_flights %>% 
      pivot_wider(names_from = 'dep_time < 801', values_from = 
                'early.bird.flights.per.carrier') %>%
      rename(Regular.Flight = 'FALSE', Early.Bird.Flight = 'TRUE', 
             Unknown = 'NA')

print(wide_early_bird_flights)
```

After this, we once again need to add in columns that list the total flights
considered and the percentage of flights classified as "early bird flights", or
before 8 AM. This is first done by adding the columns 'Regular.Flight' and 
'Early.Bird.Flight', giving us the column and values for 'Total.Flights'. Next,
we take the 'Early.Bird.Flight' and divide it by 'Total.Flights', then 
multiplying it by 100. This gives us the percentage by carrier of flights that 
were before 8 AM.

```{r}
early_bird_flight_percentage <- wide_early_bird_flights %>%
  mutate(Total.Flights = Early.Bird.Flight + Regular.Flight) %>%
  mutate(Early.Bird.Flight.Percentage = (Early.Bird.Flight / Total.Flights)*100)

print(early_bird_flight_percentage)
```

Lastly, we can once again graph our data by carrier and percentage of flights 
before 8 AM. This is done via the commands below.

```{r}
early_bird_plot <- ggplot(early_bird_flight_percentage, 
                               aes(x = carrier, y =  Early.Bird.Flight.Percentage))+
                  ggtitle(label= "Early-Bird Flights by Carrier",
                          subtitle= "Percentage of Flights Departing Before 8 AM")+
                  geom_bar(stat = "identity")

print(early_bird_plot)
```
As you can see, the percentage of flights that take off before 8 AM varied 
drastically by carrier. 

The entire code for this set of data is listed below.

```{r}
library(tidyverse)
library(nycflights13)

early_bird_flights <- flights %>%
  group_by(dep_time < 801, carrier) %>%
  summarise(early.bird.flights.per.carrier = n(), .groups='keep')


wide_early_bird_flights <- early_bird_flights %>% 
      pivot_wider(names_from = 'dep_time < 801', values_from = 
                'early.bird.flights.per.carrier') %>%
      rename(Regular.Flight = 'FALSE', Early.Bird.Flight = 'TRUE', 
             Unknown = 'NA')

early_bird_flight_percentage <- wide_early_bird_flights %>%
  mutate(Total.Flights = Early.Bird.Flight + Regular.Flight) %>%
  mutate(Early.Bird.Flight.Percentage = (Early.Bird.Flight / Total.Flights)*100)

early_bird_plot <- ggplot(early_bird_flight_percentage, 
                               aes(x = carrier, y =  Early.Bird.Flight.Percentage))+
                  ggtitle(label= "Early-Bird Flights by Carrier",
                          subtitle= "Percentage of Flights Departing Before 8 AM")+
                  geom_bar(stat = "identity")
```

